Easy as Pi, Volume II: On the Goodput and Energy of Network Coding Schemes in the Raspberry Pi

### 2016-08-15 measurements.tex
more changes

%With the testbed, methodology and setups from the previous sections, we
With the methodology and setups from the previous sections, we
proceed to obtain the measurements for the \ac{Raspi} 1 and 2 devices.
We consider the following set of parameters for our study:
For all the codes, we use $g = [16,\ 32,\ 64,\ 128,\ 256,\ 512]$
and $q = [2,\ 2^8]$. For the single-core implementations
and cases when the generation size is varied $B = 1600$ bytes.
We consider a setup where we only vary the packet size,
$B = [64,\ 128,\ 256,\ 512,\ 1024,\ 2048,\ 4096,\ 8192,\ 16384,\ 32768
,\ 65536,\ 131072]$ bytes with a generation size fixed on $g = [16,\ 128]$
to see the performance of the \ac{Raspi}s in low and high
packet size regimes. We considered $B = 1536$ bytes for the optimized
multicore implementation. For \ac{SRLNC}, we considered
$d = [0.02,\ 0.1]$. For \ac{TSNC}, we considered excess packets,
$e = [8,\ 16]$, so that the budget is $b = g + e$. In all our
measurement reports, to simplify their review, we first present the
results for the \ac{Raspi} 1 and later continue with the \ac{Raspi} 2.

\subsection{Goodput}
For the goodput, we separate the results according to their time benchmarks
as we did in Section \ref{sec:metrics} since their consider different
methodologies. We proceed first with the measurements related to the encoder
and later review the case of the decoder.

@@

Similarly, Fig. \ref{fig:enc_good_rasp2} shows the same set of
measurements for the \ac{Raspi} 2. Hence, Fig.
\ref{fig:enc_good_rasp2_gen_gf2} and \ref{fig:enc_good_rasp2_gen_gf256}
present goodput as function of the generation size  and
Fig. \ref{fig:enc_good_rasp2_packet_gf2} and
\ref{fig:enc_good_rasp2_packet_gf256} as a function of the packet size
for the same cases mentioned previously. Therefore, we will proceed
to make the results analysis with \ac{Raspi} 1 and indicate which
similarities or differences occur for the \ac{Raspi} 2 and when do
they occur.
%
\begin{figure*}
    \centering
    \begin{subfigure}[b]{0.475\textwidth}
        \centering
        \includegraphics[width=1.15\textwidth]{images/06_06_2016/goodput_vs_generation_size_Rasp_encoder_Binary_1600.eps}
        \caption[]%
        {{\small Goodput vs. Generation size for $q = 2$}}
        \label{fig:enc_good_rasp1_gen_gf2}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.475\textwidth}
        \centering
        \includegraphics[width=1.15\textwidth]{images/06_06_2016/goodput_vs_generation_size_Rasp_encoder_Binary8_1600.eps}
        \caption[]%
        {{\small Goodput vs. Generation size for $q = 2^8$}}
        \label{fig:enc_good_rasp1_gen_gf256}
    \end{subfigure}
    \vskip\baselineskip
    \begin{subfigure}[b]{0.475\textwidth}
        \centering
        \includegraphics[width=1.15\textwidth]{images/06_06_2016/goodput_vs_symbol_size_Rasp_encoder_Binary_16.eps}
        \caption[]%
        {{\small Goodput vs. Packet size for $q = 2,\ g=16$}}
        \label{fig:enc_good_rasp1_packet_gf2}
    \end{subfigure}
    \quad
    \begin{subfigure}[b]{0.475\textwidth}
        \centering
        \includegraphics[width=1.15\textwidth]{images/06_06_2016/goodput_vs_symbol_size_Rasp_encoder_Binary8_16.eps}
        \caption[]%
        {{\small Goodput vs. Packet size for $q = 2^8,\ g=16$}}
        \label{fig:enc_good_rasp1_packet_gf256}
    \end{subfigure}
    \caption[]
    {\small Encoder goodput measurements for the \ac{Raspi} 1}
    \label{fig:enc_good_rasp1}
\end{figure*}
%
As it can seen in the from: Fig. \ref{fig:enc_good_rasp1_gen_gf2},
\ref{fig:enc_good_rasp1_gen_gf256}, \ref{fig:enc_good_rasp2_gen_gf2} and
\ref{fig:enc_good_rasp2_gen_gf256}; which indicate the goodput dependency
on the generation size, the encoding goodput gets reduced as the generation
size increases, regardless of the \ac{Raspi} model observed. The reason
is that the encoding operation processing is $\mathcal{O}(g)$ because
it is required to create $g$ coding coefficients and do $g$ multiplications
for each packet. This makes the goodput to scale as the inverse of the
generation size.

In the sets of figures shown in Fig.~\ref{fig:enc_good_rasp1} and
\ref{fig:enc_good_rasp2}, it can be observed that the goodput for $GF(2)$
is higher than for $GF(2^8)$ but still around the same order of magnitude.
This difference is explained by noticing that \ac{GF} arithmetics in the
binary field are simply XOR or AND operations. These operations are
implemented efficiently by most of the architectures nowadays and are
tailored to be fast in principle. However, the operations in $GF(2^8)$
are more complex given that the finite field arithmetics have to be
performed with lookup tables which at the end reduces the computing
speed giving a lower goodput when compared with the binary field.

In Fig.~\ref{fig:enc_good_rasp1_gen_gf2}, \ref{fig:enc_good_rasp1_gen_gf256},
\ref{fig:enc_good_rasp2_gen_gf2} and \ref{fig:enc_good_rasp2_gen_gf256};
it can be seen that the goodput trends of five codes as a function of
the generation size: \ac{RLNC}, \ac{SRLNC} with $d = [0.02,\ 0.1]$
and \ac{TSNC} with an extra parameter that we mention. We define
$C$ as the number of density regions in the \ac{TSNC} transmission
process. The maximum number of possible regions depends
on the generation size. See~\eqref{eq:feedback_dofs}.
The larger the generation, the more density regions may
be formed and more density changes are possible.
Throughout this paper, \ac{TSNC} is configured to use the maximum number
of density regions possible. 
As there can be at least $C=1$ density region in a transmission, we use
$C=0$ to indicate the maximum possible density regions. This is
used for \ac{TSNC} in plots where the generation size is not fixed.
%In fact, \ac{TSNC} is configured to use the
%maximum possible density regions in all plots throughout this paper.
%The larger the generation, the more regions are formed and
%more density changes are possible.
%Throughout the section, we evaluate \ac{TSNC} with as many possible
%density regions that are possible to form during a transmission. This number
%grows as the generation increases, so we define $C=0$ as the maximum
%number of possible regions.
%
%We first evaluate \ac{TSNC} with no
%changes, e.g. $C = 0$ and the two amounts for dependent packets mentioned
%earlier. This means that we tolerate at most 8 or 16\fxnote{dependent packets} packets regardless of
%the generation. The goal with setting first \ac{TSNC} with these values
%is to observe first the sole effect of the budget in the code since
%$C = 0$ implies no density changes during the whole process.
%
For the plots in the mentioned figures from the encoder goodput measurements,
\ac{RLNC} presents the lowest performance in terms of goodput and
\ac{TSNC} with $e = 16$, the highest regardless of the \ac{Raspi} model.
Given that the processing time depends on the amount of coded packets
required to create, \ac{RLNC} is the slowest to process since it must
use all of the $g$ original packets
%\fxnote{The encoding process speed depends on the total number of original packets to mix. Thus coding density and number of coded packets per generation}.
Later, sparse codes process the data
at a larger rate since less packets are being mixed when creating a
coded packet. The caveat of these schemes is that the sparser the code,
the more probable the ocurrences of linearly dependent packets is. So, basically
the sparser the codes, then more overhead due to transmissions of linearly dependent
packets since it is more frequent to generate similar set of packets.
Excluding the coding coefficients overhead, the overhead due to
transmissions of linearly dependent packets might be high for the sparse schemes.
For example, if we consider \ac{TSNC} with $e = 16$ and $g = 16$
in Fig.~\ref{fig:enc_good_rasp1_gen_gf2}, the budget in this case permits
to send up to 32 packets which is $2\times$ the generation
size for an overhead of $100\%$ excluding the overhead from appending
the coding coefficients. This happens because \ac{TSNC}
has been allowed to add too much redundancy in this case. Given that $C=0$,
this permits it to use a single very low code density $d$. For \ac{RLNC},
this is not the case since the occurrence of linearly dependent coded packets is
small, because all coding coefficients are used. Even for $GF(2)$, the
average amount of redundant packets for \ac{RLNC} has proven to be 1.6
packets after $g$ have been transmitted \cite{trullols2011exact,zhao2012notes},
but less than the cases where sparse codes are utilized. Overall, we
observe that there is a trade-off between goodput and linearly dependent coded
packets transmission overhead.

@@

If we consider the previous effect when reviewing the trends in the
mentioned figures, in all the models we observe that the maximum is not
at 16 KiB but at a smaller value depending on the model, field size
and coding scheme considered. The reason is that the \ac{CPU} needs to
allocate computational resources to other processes from the different tasks
running in the \ac{Raspi}. Then, given there are various
tasks from other applications for proper functioning running in the
\ac{Raspi} at the same time as coding, the cache space is filled also
with the data from these tasks, thus diminishing the cache space available
for coding operations.

For the \ac{Raspi} 1 model, we observe in
Fig.~\ref{fig:enc_good_rasp1_packet_gf2} that the critical
packet size for \ac{RLNC} using $GF(2)$ occurs at 1 KiB whereas for the
sparse codes is close to 8 KiB in the most of the cases for $GF(2)$. This
difference takes place since the sparse codes mix less packets their
process than \ac{RLNC}, which turns into less data loading in the
cache for doing computations. For a density of $d = 0.1$, packet size of
$B = 8$ KiB and $g = 16$ packets, we observe that roughly $\ceil{gd} = 2$
packets are inserted in the cache when calculating a coded packet with this
sparse code. Loading this into the cache, stands for 16 KiB which is the
cache size. A similar effect occurs for the other sparse codes. However, for
\ac{RLNC} given that it is a dense code since $d \to 1$, \ac{RLNC}
packets load data from all the $g = 16$ coding coefficients, which accounts
for the 16 KiB of the cache size. In
Fig.~\ref{fig:enc_good_rasp1_packet_gf256}, although the ideal packet
size remains the same for \ac{RLNC} and the sparse codes, the final goodput
is lower due to the field size effect.
%
\begin{figure*}
    \centering
    \begin{subfigure}[b]{0.475\textwidth}
        \centering
        \includegraphics[width=1.15\textwidth]{images/06_06_2016/goodput_vs_generation_size_Rasp_v2_encoder_Binary_1600.eps}
        \caption[]%
        {{\small Goodput vs. Generation size for $q = 2$}}
        \label{fig:enc_good_rasp2_gen_gf2}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.475\textwidth}
        \centering
        \includegraphics[width=1.15\textwidth]{images/06_06_2016/goodput_vs_generation_size_Rasp_v2_encoder_Binary8_1600.eps}
        \caption[]%
        {{\small Goodput vs. Generation size for $q = 2^8$}}
        \label{fig:enc_good_rasp2_gen_gf256}
    \end{subfigure}
    \vskip\baselineskip
    \begin{subfigure}[b]{0.475\textwidth}
        \centering
        \includegraphics[width=1.15\textwidth]{images/06_06_2016/goodput_vs_symbol_size_Rasp_v2_encoder_Binary_16.eps}
        \caption[]%
        {{\small Goodput vs. Packet size for $q = 2,\ g=16$}}
        \label{fig:enc_good_rasp2_packet_gf2}
    \end{subfigure}
    \quad
    \begin{subfigure}[b]{0.475\textwidth}
        \centering
        \includegraphics[width=1.15\textwidth]{images/06_06_2016/goodput_vs_symbol_size_Rasp_v2_encoder_Binary8_16.eps}
        \caption[]%
        {{\small Goodput vs. Packet size for $q = 2^8,\ g=16$}}
        \label{fig:enc_good_rasp2_packet_gf256}
    \end{subfigure}
    \caption[]
    {\small Encoder goodput measurements for the \ac{Raspi} 2}
    \label{fig:enc_good_rasp2}
\end{figure*}
%
\par
The effects mentioned for the \ac{Raspi} 1 were also observed for
the \ac{Raspi} 2 as mentioned previously. Still, the \ac{Raspi} 2 achieves
roughly $5\times$ to $7\times$ gains in terms of encoding speed when comparing
the goodputs in Fig.~\ref{fig:enc_good_rasp1} and \ref{fig:enc_good_rasp2},
given that it has an \ac{ARM} Cortex A7 (v7) \ac{CPU} and twice the \ac{RAM}
size than the \ac{ARM}1176JZF-S (v6) core of the \ac{Raspi} 1 model.

>>>>

%With the testbed, methodology and setups from the previous sections, we
With the methodology and setups from the previous sections, we
proceed to obtain the measurements for the \ac{Raspi} 1 and 2 devices.
We consider the following set of parameters for our study:
For all the codes, we use $g = [16,\ 32,\ 64,\ 128,\ 256,\ 512]$
and $q = [2,\ 2^8]$. For the single-core implementations
and cases when the generation size is varied $B = 1600$ bytes.
We consider another setup where only the packet size varies,
$B = [64,\ 128,\ 256,\ 512,\ 1024,\ 2048,\ 4096,\ 8192,\ 16384,\ 32768
,\ 65536,\ 131072]$ bytes with a generation size fixed on $g = [16,\ 128]$
to see the performance of the \ac{Raspi}s in low and high
packet size regimes. The third setup we considered used $B = 1536$ bytes for the optimized
multicore implementation. 
%For \ac{SRLNC}, we considered $d = [0.02,\ 0.1]$.
\ac{SRLNC} was measured with the densities $d = [0.02,\ 0.1]$.
For \ac{TSNC}, we considered excess packets,
$e = [8,\ 16]$, so that the budget is $b = g + e$. In all our
measurement reports, to simplify their review, we first present the
results for the \ac{Raspi} 1 and later continue with the \ac{Raspi} 2.

\subsection{Goodput}
For the goodput, we separate the results according to their time benchmarks
as we did in Section \ref{sec:metrics}.
%since their consider different methodologies.
We proceed first with the measurements related to the encoder
and later review the case of the decoder.

@@

Similarly, Fig. \ref{fig:enc_good_rasp2} shows the same set of
measurements for the \ac{Raspi} 2. Hence, Fig.
\ref{fig:enc_good_rasp2_gen_gf2} and \ref{fig:enc_good_rasp2_gen_gf256}
present goodput as function of the generation size  and
Fig. \ref{fig:enc_good_rasp2_packet_gf2} and
\ref{fig:enc_good_rasp2_packet_gf256} as a function of the packet size
for the same cases mentioned previously. Therefore, we will proceed
to make the results analysis with \ac{Raspi} 1 and indicate which
similarities or differences occur for the \ac{Raspi} 2 and when do
they occur.
%
\begin{figure*}
    \centering
    \begin{subfigure}[b]{0.475\textwidth}
        \centering
        \includegraphics[width=1.15\textwidth]{images/06_06_2016/goodput_vs_generation_size_Rasp_encoder_Binary_1600.eps}
        \caption[]%
        {{\small Goodput vs. Generation size for $q = 2$}}
        \label{fig:enc_good_rasp1_gen_gf2}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.475\textwidth}
        \centering
        \includegraphics[width=1.15\textwidth]{images/06_06_2016/goodput_vs_generation_size_Rasp_encoder_Binary8_1600.eps}
        \caption[]%
        {{\small Goodput vs. Generation size for $q = 2^8$}}
        \label{fig:enc_good_rasp1_gen_gf256}
    \end{subfigure}
    \vskip\baselineskip
    \begin{subfigure}[b]{0.475\textwidth}
        \centering
        \includegraphics[width=1.15\textwidth]{images/06_06_2016/goodput_vs_symbol_size_Rasp_encoder_Binary_16.eps}
        \caption[]%
        {{\small Goodput vs. Packet size for $q = 2,\ g=16$}}
        \label{fig:enc_good_rasp1_packet_gf2}
    \end{subfigure}
    \quad
    \begin{subfigure}[b]{0.475\textwidth}
        \centering
        \includegraphics[width=1.15\textwidth]{images/06_06_2016/goodput_vs_symbol_size_Rasp_encoder_Binary8_16.eps}
        \caption[]%
        {{\small Goodput vs. Packet size for $q = 2^8,\ g=16$}}
        \label{fig:enc_good_rasp1_packet_gf256}
    \end{subfigure}
    \caption[]
    {\small Encoder goodput measurements for the \ac{Raspi} 1}
    \label{fig:enc_good_rasp1}
\end{figure*}
%
As it can seen from Fig. \ref{fig:enc_good_rasp1_gen_gf2},
\ref{fig:enc_good_rasp1_gen_gf256}, \ref{fig:enc_good_rasp2_gen_gf2} and
\ref{fig:enc_good_rasp2_gen_gf256}; which indicate the goodput dependency
on the generation size, the encoding goodput gets reduced as the generation
size increases, regardless of the \ac{Raspi} model observed. The reason
is that the encoding operation processing is $\mathcal{O}(g)$ because
it is required to create $g$ coding coefficients and do $g$ multiplications
for each packet. This makes the goodput scale as the inverse of the
generation size.

In the sets of figures shown in Fig.~\ref{fig:enc_good_rasp1} and
\ref{fig:enc_good_rasp2}, it can be observed that the goodput for $GF(2)$
is higher than for $GF(2^8)$ but still around the same order of magnitude.
This difference is explained by noticing that \ac{GF} arithmetics in the
binary field are simply XOR or AND operations. These operations are
implemented efficiently in nowadays architectures.
%and are tailored to be fast in principle.
However, the operations in $GF(2^8)$
are more complex given that the finite field arithmetics have to be
performed with lookup tables which at the end reduces the computing
speed giving a lower goodput when compared with the binary field.

In Fig.~\ref{fig:enc_good_rasp1_gen_gf2}, \ref{fig:enc_good_rasp1_gen_gf256},
\ref{fig:enc_good_rasp2_gen_gf2} and \ref{fig:enc_good_rasp2_gen_gf256};
it can be seen that the goodput trends of five codes as a function of
the generation size: \ac{RLNC}, \ac{SRLNC} with $d = [0.02,\ 0.1]$
and \ac{TSNC} with an extra parameter that we mention. We define
$C$ as the number of density regions in the \ac{TSNC} transmission
process. The maximum number of possible regions depends
on the generation size~\eqref{eq:feedback_dofs}.
The larger the generation, the more density regions may
be formed and more density changes are possible.
Throughout this paper, \ac{TSNC} is configured to use the maximum number
of density regions possible. 
As there can be at least $C=1$ density region in a transmission, we use
$C=0$ to indicate the maximum possible density regions. This is
used for \ac{TSNC} in plots where the generation size is not fixed.
%In fact, \ac{TSNC} is configured to use the
%maximum possible density regions in all plots throughout this paper.
%The larger the generation, the more regions are formed and
%more density changes are possible.
%Throughout the section, we evaluate \ac{TSNC} with as many possible
%density regions that are possible to form during a transmission. This number
%grows as the generation increases, so we define $C=0$ as the maximum
%number of possible regions.
%
%We first evaluate \ac{TSNC} with no
%changes, e.g. $C = 0$ and the two amounts for dependent packets mentioned
%earlier. This means that we tolerate at most 8 or 16\fxnote{dependent packets} packets regardless of
%the generation. The goal with setting first \ac{TSNC} with these values
%is to observe first the sole effect of the budget in the code since
%$C = 0$ implies no density changes during the whole process.
%
For the plots in the mentioned figures from the encoder goodput measurements,
\ac{RLNC} presents the lowest performance in terms of goodput and
\ac{TSNC} with $e = 16$, the highest regardless of the \ac{Raspi} model.
Given that the processing time depends on the amount of coded packets
required to create, \ac{RLNC} is the slowest to process since it must
use all of the $g$ original packets.
%\fxnote{The encoding process speed depends on the total number of original packets to mix. Thus coding density and number of coded packets per generation}.
Later, sparse codes process the data
at a larger rate since less packets are being mixed when creating a
coded packet. The caveat of these schemes is that the sparser the code,
the more probable the ocurrences of linearly dependent packets is. So, basically
the sparser the codes, then more overhead due to transmissions of linearly dependent
packets.
%since it is more frequent to generate similar set of packets.
Excluding the coding coefficients overhead, the overhead due to
transmissions of linearly dependent packets might be high for the sparse schemes.
For example, if we consider \ac{TSNC} with $e = 16$ and $g = 16$
in Fig.~\ref{fig:enc_good_rasp1_gen_gf2}, the budget in this case permits
to send up to 32 packets which is $2\times$ the generation
size for an overhead of $100\%$ excluding the overhead from appending
the coding coefficients. This happens because \ac{TSNC}
has been allowed to add too much redundancy in this case.
%Given that $C=0$, this permits it to use a single very low code density $d$.
For \ac{RLNC},
this is not the case since the occurrence of linearly dependent coded packets is
small, because all coding coefficients are used. Even for $GF(2)$, the
average amount of redundant packets for \ac{RLNC} has proven to be 1.6
packets after $g$ have been transmitted \cite{trullols2011exact,zhao2012notes},
but less than the cases where sparse codes are utilized. Overall, we
observe that there is a trade-off between goodput and linearly dependent coded
packets transmission overhead.

@@

If we consider the previous effect when reviewing the trends in the
mentioned figures, in all the models, we observe that the maximum coding performance is not
at 16 KiB but at a smaller value depending on the model, field size
and coding scheme considered. The reason is that the \ac{CPU} needs to
allocate computational resources to other processes from the different tasks
running in the \ac{Raspi}. Then, given there are various
tasks from other applications for proper functioning running in the
\ac{Raspi} at the same time as coding, the cache space is filled also
with the data from these tasks, thus diminishing the cache space available
for coding operations.

For the \ac{Raspi} 1 model, we observe in
Fig.~\ref{fig:enc_good_rasp1_packet_gf2} that the critical
packet size for \ac{RLNC} using $GF(2)$ occurs at 1 KiB whereas for the
sparse codes is close to 8 KiB in most cases for $GF(2)$. This
difference takes place since the sparse codes mix less packets
than \ac{RLNC}, which turns into less data loading in the
cache for doing computations. For a density of $d = 0.1$, packet size of
$B = 8$ KiB and $g = 16$ packets, we observe that roughly $\ceil{gd} = 2$
packets are inserted in the cache when calculating a coded packet with this
sparse code. Loading this into the cache, stands for 16 KiB which is the
cache size. A similar effect occurs for the other sparse codes. However, for
\ac{RLNC} given that it is a dense code since $d \to 1$, \ac{RLNC}
packets load data from all the $g = 16$ coding coefficients, which accounts
for the 16 KiB of the cache size. In
Fig.~\ref{fig:enc_good_rasp1_packet_gf256}, although the ideal packet
size remains the same for \ac{RLNC} and the sparse codes, the final goodput
is lower due to the field size effect.
%
\begin{figure*}
    \centering
    \begin{subfigure}[b]{0.475\textwidth}
        \centering
        \includegraphics[width=1.15\textwidth]{images/06_06_2016/goodput_vs_generation_size_Rasp_v2_encoder_Binary_1600.eps}
        \caption[]%
        {{\small Goodput vs. Generation size for $q = 2$}}
        \label{fig:enc_good_rasp2_gen_gf2}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.475\textwidth}
        \centering
        \includegraphics[width=1.15\textwidth]{images/06_06_2016/goodput_vs_generation_size_Rasp_v2_encoder_Binary8_1600.eps}
        \caption[]%
        {{\small Goodput vs. Generation size for $q = 2^8$}}
        \label{fig:enc_good_rasp2_gen_gf256}
    \end{subfigure}
    \vskip\baselineskip
    \begin{subfigure}[b]{0.475\textwidth}
        \centering
        \includegraphics[width=1.15\textwidth]{images/06_06_2016/goodput_vs_symbol_size_Rasp_v2_encoder_Binary_16.eps}
        \caption[]%
        {{\small Goodput vs. Packet size for $q = 2,\ g=16$}}
        \label{fig:enc_good_rasp2_packet_gf2}
    \end{subfigure}
    \quad
    \begin{subfigure}[b]{0.475\textwidth}
        \centering
        \includegraphics[width=1.15\textwidth]{images/06_06_2016/goodput_vs_symbol_size_Rasp_v2_encoder_Binary8_16.eps}
        \caption[]%
        {{\small Goodput vs. Packet size for $q = 2^8,\ g=16$}}
        \label{fig:enc_good_rasp2_packet_gf256}
    \end{subfigure}
    \caption[]
    {\small Encoder goodput measurements for the \ac{Raspi} 2}
    \label{fig:enc_good_rasp2}
\end{figure*}
%
\par
The effects mentioned for the \ac{Raspi} 1 were also observed for
the \ac{Raspi} 2 as mentioned previously. Still, the \ac{Raspi} 2 achieves
roughly $5\times$ to $7\times$ gains in terms of encoding speed when comparing
the goodputs in Fig.~\ref{fig:enc_good_rasp1} and \ref{fig:enc_good_rasp2},
given that it has an \ac{ARM} Cortex A7 (v7) \ac{CPU} and twice the \ac{RAM}
size than the \ac{ARM}1176JZF-S (v6) core of the \ac{Raspi} 1 model.

### 2016-08-15 schemes.tex
more changes

\subsubsection{Method 1: Fixing the Coding Density}
A way to control the amount of non-zero coding coefficients, is to set a
fixed ratio of non-zero coefficients in the encoding vector of size $g$. We
refer to this fixed ratio as the average coding density $d$. Thus, for any
coded packet $C_i$ with coding coefficients $v_{ij},\ j \in [1,g]$, its
coding density is defined as follows:
%
\begin{align}
\label{eq:coding_density}
d = \frac{\sum_{j=1}^{g} f(v_{ij})}{g},\ \
    f(v_{ij}) =
    \begin{cases}
        0 &,\ v_{ij} = 0 \\
        1 &,\ v_{ij} \neq 0
    \end{cases}
\end{align}
%
From the density definition in~\eqref{eq:coding_density}, it can be
observed that $0 \leq d \leq 1$. As $g$ increased, we obtain more
granularity in the density. Notice that the special case of  $d = 0$ has no
practical purpose since it implies just zero padded data.
\textcolor{red}{exclude the case d=1 for gf(2) here}
So, a practical
range for the density excludes the zero case, $0 < d \leq 1$.
%Moreover,
%it can be also observed that the case of $d = 1$ reduces to the \ac{RLNC}
%coding scheme described in Section~\ref{ssec:RLNC}.
Moreover, the densities $d=0.5$ in $GF(2)$ and $d=1$ in higher order
\ac{GF}s is similar to the \ac{RLNC} scheme described in Section~\ref{ssec:RLNC}.

>>>>

\subsubsection{Method 1: Fixing the Coding Density}
A way to control the amount of non-zero coding coefficients, is to set a
fixed ratio of non-zero coefficients in the encoding vector of size $g$. We
refer to this fixed ratio as the average coding density $d$. Thus, for any
coded packet $C_i$ with coding coefficients $v_{ij},\ j \in [1,g]$, its
coding density is defined as follows:
%
\begin{align}
\label{eq:coding_density}
d = \frac{\sum_{j=1}^{g} f(v_{ij})}{g},\ \
    f(v_{ij}) =
    \begin{cases}
        0 &,\ v_{ij} = 0 \\
        1 &,\ v_{ij} \neq 0
    \end{cases}
\end{align}
%
From the density definition in~\eqref{eq:coding_density}, it can be
observed that $0 \leq d \leq 1$. As $g$ increased, we obtain more
granularity in the density. Notice that the special case of  $d = 0$ has no
practical purpose since it implies just zero padded data.
%In $GF(2)$ only, we also eliminate the case of $d = 1$ as it would always
%produce a coding vector of all ones.
%In $GF(2)$ only, $d = 1$ causes coding vectors to always be all ones.
%Furthermore, $d = 1$ in $GF(2)$ leads to coding vectors that are always filled
%with ones, producing the same coded packet.
%\textcolor{red}{exclude the case d=1 for gf(2) here}
So, a practical
range for the density excludes the zero case, $0 < d \leq 1$.
%range for the density that we use is $0 < d \leq 1$, and for $GF(2)$ only, $0 < d < 1$.
%Moreover,
%it can be also observed that the case of $d = 1$ reduces to the \ac{RLNC}
%coding scheme described in Section~\ref{ssec:RLNC}.
%
Furthermore, in $GF(2)$ there is no benefit of using $d > 0.5$ in terms of
generating linearly independent packets~\cite{realworld_nc2013}. Thus,
we limit the range to $0 < d \leq 0.5$ in $GF(2)$.

%Moreover, the densities $d=0.5$ in $GF(2)$ and $d=1$ in higher order
%\ac{GF}s is similar to the \ac{RLNC} scheme described in Section~\ref{ssec:RLNC}.
